Capstone Movielens Report

### 2025-05-10 reports/capstone-movielens-report.draft4.Rmd
Draft report development:  function `model.tune.param_range` (in progress)

###### Parameters
\
  - **param_values:** A list of values to search for the value corresponding to the minimum value of the `RMSE` ; 
  - **fn_tune.test.param_value:** A helper function that calculates the value of the `RMSE` for a given parameter value.; 
  - **break.if_min = TRUE:** A Boolean parameter that determines whether the function should terminate after completing the number of steps specified by the parameter `steps.beyond_min`, after the minimum value of the `RMSE` has been found;
  - **steps.beyond_min = 2:** (takes effect only if `break.if_min` parameter is `TRUE`) Specifies the number of steps after finding the minimum value of the `RMSE`, upon completion of which the function should terminate.

@@

@@

  param_values.best_result <- c(param.best_value = param.best_value, 
                                best_RMSE = best_RMSE)

  repeat{
    seq_increment <- (seq_end - seq_start)/range_divider

    if (seq_increment < 0.0000000000001) {
      warning("Function `model.tune.param_range`:
parameter value increment is too small.")

      put_log2("Function `model.tune.param_range`:
Final best RMSE for `parameter value = %1`: %2",
               param_values.best_result["param.best_value"],
               param_values.best_result["best_RMSE"])
      break
    }

@@

    file_name_tmp <- cache_file_base_name |>
      str_c("_") |>
      str_c(as.character(loop_starter[1])) |>
      str_c("_") |>
      str_c(as.character(loop_starter[3])) |>
      str_c("_") |>
      str_c(as.character(range_divider)) |>
      str_c(".") |>
      str_c(as.character(seq_start)) |>
      str_c("-") |>
      str_c(as.character(seq_end)) |>
      str_c(".RData")

    file_path_tmp <- file.path(tune_dir_path, file_name_tmp)

    put_log1("Function `model.tune.param_range`:
File path generated: %1", file_path_tmp)

    if (file.exists(file_path_tmp)) {
      put_log1("Function `model.tune.param_range`:
Loading tuning data from file: %1...", file_path_tmp)

      start <- put_start_date()
      load(file_path_tmp)
      put_end_date(start)
      put_log1("Function `model.tune.param_range`:
Tuning data has been loaded from file: %1", file_path_tmp)

      tuned.result <- tuned_result$tuned.result

      if(length(file_path_tmp) > 0) {
      }
    } else {
      tuned_result <- tune.model_param(test_param_vals, 
                                        fn_tune.test.param_value,
                                        break.if_min,
                                        steps.beyond_min)

      tuned.result <- tuned_result$tuned.result

      save(tuned_result,
           param.best_value,
           seq_increment,
           range_divider,
           file = file_path_tmp)

      put_log1("Function `model.tune.param_range`:
File saved: %1", file_path_tmp)
    }

    plot(tuned.result$parameter.value, tuned.result$RMSE)

@@

        param_values.best_result <-
          get_best_param.result(tuned.result$parameter.value,
                                tuned.result$RMSE)

        put_log2("Function `model.tune.param_range`:
      Reached the best RMSE for `parameter value = %1`: %2",
                 param_values.best_result["param.best_value"],
                 param_values.best_result["best_RMSE"])
        break
      }

@@

    param_values.best_result <- 
      get_best_param.result(tuned.result$parameter.value, 
                          tuned.result$RMSE)

    put_log2("Function `model.tune.param_range`:
Currently reached best RMSE for `parameter value = %1`: %2",
             param_values.best_result["param.best_value"],
             param_values.best_result["best_RMSE"])
  }

>>>>

###### Parameters
\

  - **param_values:** A list of values to search for the value corresponding to the minimum value of the `RMSE` ; 
  - **fn_tune.test.param_value:** A helper function that calculates the value of the `RMSE` for a given parameter value.; 
  - **break.if_min = TRUE:** A Boolean parameter that determines whether the function should terminate after completing the number of steps specified by the parameter `steps.beyond_min`, after the minimum value of the `RMSE` has been found;
  - **steps.beyond_min = 2:** (takes effect only if `break.if_min` parameter is `TRUE`) Specifies the number of steps after finding the minimum value of the `RMSE`, upon completion of which the function should terminate.

@@

###### Parameters
\

  - **fn_tune.test.param_value:** A helper function that calculates the value of the `RMSE` for a given parameter value.; 
  - **break.if_min = TRUE:** A Boolean parameter that determines whether the function should terminate after completing the number of steps specified by the parameter `steps.beyond_min`, after the minimum value of the `RMSE` has been found;
  - **steps.beyond_min = 2:** (takes effect only if `break.if_min` parameter is `TRUE`) Specifies the number of steps after finding the minimum value of the `RMSE`, upon completion of which the function should terminate.

@@

  param_values.best_result <- c(param.best_value = param.best_value, 
                                best_RMSE = best_RMSE)
  # Start repeat loop
  repeat{
    seq_increment <- (seq_end - seq_start)/range_divider

    if (seq_increment < 0.0000000000001) {
      warning("Function `model.tune.param_range`:
parameter value increment is too small.")
      break
    }

@@

    tuned_result <- tune.model_param(test_param_vals, 
                                  fn_tune.test.param_value,
                                  break.if_min,
                                  steps.beyond_min)

    tuned.result <- tuned_result$tuned.result
    plot(tuned.result$parameter.value, tuned.result$RMSE)

@@

        param_values.best_result <-
          get_best_param.result(tuned.result$parameter.value,
                                tuned.result$RMSE)
        break
      }

@@

    param_values.best_result <- 
      get_best_param.result(tuned.result$parameter.value, 
                          tuned.result$RMSE)
  }
  # End repeat loop

### 2025-05-10 reports/capstone-movielens-report.draft4.tex
Draft report development:  function `model.tune.param_range` (in progress)

\hfill\break
- \textbf{param\_values:} A list of values to search for the value
corresponding to the minimum value of the \texttt{RMSE} ; -
\textbf{fn\_tune.test.param\_value:} A helper function that calculates
the value of the \texttt{RMSE} for a given parameter value.; -
\textbf{break.if\_min = TRUE:} A Boolean parameter that determines
whether the function should terminate after completing the number of
steps specified by the parameter \texttt{steps.beyond\_min}, after the
minimum value of the \texttt{RMSE} has been found; -
\textbf{steps.beyond\_min = 2:} (takes effect only if
\texttt{break.if\_min} parameter is \texttt{TRUE}) Specifies the number
of steps after finding the minimum value of the \texttt{RMSE}, upon
completion of which the function should terminate.

@@

@@

\NormalTok{  param\_values.best\_result }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{param.best\_value =}\NormalTok{ param.best\_value, }
                                \AttributeTok{best\_RMSE =}\NormalTok{ best\_RMSE)}

  \ControlFlowTok{repeat}\NormalTok{\{}
\NormalTok{    seq\_increment }\OtherTok{\textless{}{-}}\NormalTok{ (seq\_end }\SpecialCharTok{{-}}\NormalTok{ seq\_start)}\SpecialCharTok{/}\NormalTok{range\_divider }

    \ControlFlowTok{if}\NormalTok{ (seq\_increment }\SpecialCharTok{\textless{}} \FloatTok{0.0000000000001}\NormalTok{) \{}
      \FunctionTok{warning}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{parameter value increment is too small."}\NormalTok{)}

      \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{Final best RMSE for \textasciigrave{}parameter value = \%1\textasciigrave{}: \%2"}\NormalTok{,}
\NormalTok{               param\_values.best\_result[}\StringTok{"param.best\_value"}\NormalTok{],}
\NormalTok{               param\_values.best\_result[}\StringTok{"best\_RMSE"}\NormalTok{])}
      \ControlFlowTok{break}
\NormalTok{    \}}

@@

\NormalTok{    file\_name\_tmp }\OtherTok{\textless{}{-}}\NormalTok{ cache\_file\_base\_name }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\StringTok{"\_"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(loop\_starter[}\DecValTok{1}\NormalTok{])) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\StringTok{"\_"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(loop\_starter[}\DecValTok{3}\NormalTok{])) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\StringTok{"\_"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(range\_divider)) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\StringTok{"."}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(seq\_start)) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\StringTok{"{-}"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(seq\_end)) }\SpecialCharTok{|\textgreater{}}
      \FunctionTok{str\_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{)}

\NormalTok{    file\_path\_tmp }\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(tune\_dir\_path, file\_name\_tmp)}

    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{File path generated: \%1"}\NormalTok{, file\_path\_tmp)}

    \ControlFlowTok{if}\NormalTok{ (}\FunctionTok{file.exists}\NormalTok{(file\_path\_tmp)) \{}
      \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{Loading tuning data from file: \%1..."}\NormalTok{, file\_path\_tmp)}

\NormalTok{      start }\OtherTok{\textless{}{-}} \FunctionTok{put\_start\_date}\NormalTok{()}
      \FunctionTok{load}\NormalTok{(file\_path\_tmp)}
      \FunctionTok{put\_end\_date}\NormalTok{(start)}
      \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{Tuning data has been loaded from file: \%1"}\NormalTok{, file\_path\_tmp)}

\NormalTok{      tuned.result }\OtherTok{\textless{}{-}}\NormalTok{ tuned\_result}\SpecialCharTok{$}\NormalTok{tuned.result}

      \ControlFlowTok{if}\NormalTok{(}\FunctionTok{length}\NormalTok{(file\_path\_tmp) }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{) \{}
\NormalTok{      \}}
\NormalTok{    \} }\ControlFlowTok{else}\NormalTok{ \{}
\NormalTok{      tuned\_result }\OtherTok{\textless{}{-}} \FunctionTok{tune.model\_param}\NormalTok{(test\_param\_vals, }
\NormalTok{                                        fn\_tune.test.param\_value,}
\NormalTok{                                        break.if\_min,}
\NormalTok{                                        steps.beyond\_min)}

\NormalTok{      tuned.result }\OtherTok{\textless{}{-}}\NormalTok{ tuned\_result}\SpecialCharTok{$}\NormalTok{tuned.result}

      \FunctionTok{save}\NormalTok{(tuned\_result,}
\NormalTok{           param.best\_value,}
\NormalTok{           seq\_increment,}
\NormalTok{           range\_divider,}
           \AttributeTok{file =}\NormalTok{ file\_path\_tmp)}

      \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{File saved: \%1"}\NormalTok{, file\_path\_tmp)}
\NormalTok{    \}}

    \FunctionTok{plot}\NormalTok{(tuned.result}\SpecialCharTok{$}\NormalTok{parameter.value, tuned.result}\SpecialCharTok{$}\NormalTok{RMSE)}

@@

\NormalTok{        param\_values.best\_result }\OtherTok{\textless{}{-}}
          \FunctionTok{get\_best\_param.result}\NormalTok{(tuned.result}\SpecialCharTok{$}\NormalTok{parameter.value,}
\NormalTok{                                tuned.result}\SpecialCharTok{$}\NormalTok{RMSE)}

        \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{      Reached the best RMSE for \textasciigrave{}parameter value = \%1\textasciigrave{}: \%2"}\NormalTok{,}
\NormalTok{                 param\_values.best\_result[}\StringTok{"param.best\_value"}\NormalTok{],}
\NormalTok{                 param\_values.best\_result[}\StringTok{"best\_RMSE"}\NormalTok{])}
        \ControlFlowTok{break}
\NormalTok{      \}}

@@

\NormalTok{    param\_values.best\_result }\OtherTok{\textless{}{-}} 
      \FunctionTok{get\_best\_param.result}\NormalTok{(tuned.result}\SpecialCharTok{$}\NormalTok{parameter.value, }
\NormalTok{                          tuned.result}\SpecialCharTok{$}\NormalTok{RMSE)}

    \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{Currently reached best RMSE for \textasciigrave{}parameter value = \%1\textasciigrave{}: \%2"}\NormalTok{,}
\NormalTok{             param\_values.best\_result[}\StringTok{"param.best\_value"}\NormalTok{],}
\NormalTok{             param\_values.best\_result[}\StringTok{"best\_RMSE"}\NormalTok{])}
\NormalTok{  \}}

>>>>

\begin{itemize}
\tightlist
\item
  \textbf{param\_values:} A list of values to search for the value
  corresponding to the minimum value of the \texttt{RMSE} ;
\item
  \textbf{fn\_tune.test.param\_value:} A helper function that calculates
  the value of the \texttt{RMSE} for a given parameter value.;
\item
  \textbf{break.if\_min = TRUE:} A Boolean parameter that determines
  whether the function should terminate after completing the number of
  steps specified by the parameter \texttt{steps.beyond\_min}, after the
  minimum value of the \texttt{RMSE} has been found;
\item
  \textbf{steps.beyond\_min = 2:} (takes effect only if
  \texttt{break.if\_min} parameter is \texttt{TRUE}) Specifies the
  number of steps after finding the minimum value of the \texttt{RMSE},
  upon completion of which the function should terminate.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  \textbf{fn\_tune.test.param\_value:} A helper function that calculates
  the value of the \texttt{RMSE} for a given parameter value.;
\item
  \textbf{break.if\_min = TRUE:} A Boolean parameter that determines
  whether the function should terminate after completing the number of
  steps specified by the parameter \texttt{steps.beyond\_min}, after the
  minimum value of the \texttt{RMSE} has been found;
\item
  \textbf{steps.beyond\_min = 2:} (takes effect only if
  \texttt{break.if\_min} parameter is \texttt{TRUE}) Specifies the
  number of steps after finding the minimum value of the \texttt{RMSE},
  upon completion of which the function should terminate.
\end{itemize}

@@

\NormalTok{  param\_values.best\_result }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{param.best\_value =}\NormalTok{ param.best\_value, }
                                \AttributeTok{best\_RMSE =}\NormalTok{ best\_RMSE)}
  \CommentTok{\# Start repeat loop}
  \ControlFlowTok{repeat}\NormalTok{\{}
\NormalTok{    seq\_increment }\OtherTok{\textless{}{-}}\NormalTok{ (seq\_end }\SpecialCharTok{{-}}\NormalTok{ seq\_start)}\SpecialCharTok{/}\NormalTok{range\_divider }

    \ControlFlowTok{if}\NormalTok{ (seq\_increment }\SpecialCharTok{\textless{}} \FloatTok{0.0000000000001}\NormalTok{) \{}
      \FunctionTok{warning}\NormalTok{(}\StringTok{"Function \textasciigrave{}model.tune.param\_range\textasciigrave{}:}
\StringTok{parameter value increment is too small."}\NormalTok{)}
      \ControlFlowTok{break}
\NormalTok{    \}}

@@

\NormalTok{    tuned\_result }\OtherTok{\textless{}{-}} \FunctionTok{tune.model\_param}\NormalTok{(test\_param\_vals, }
\NormalTok{                                  fn\_tune.test.param\_value,}
\NormalTok{                                  break.if\_min,}
\NormalTok{                                  steps.beyond\_min)}

\NormalTok{    tuned.result }\OtherTok{\textless{}{-}}\NormalTok{ tuned\_result}\SpecialCharTok{$}\NormalTok{tuned.result}
    \FunctionTok{plot}\NormalTok{(tuned.result}\SpecialCharTok{$}\NormalTok{parameter.value, tuned.result}\SpecialCharTok{$}\NormalTok{RMSE)}

@@

\NormalTok{        param\_values.best\_result }\OtherTok{\textless{}{-}}
          \FunctionTok{get\_best\_param.result}\NormalTok{(tuned.result}\SpecialCharTok{$}\NormalTok{parameter.value,}
\NormalTok{                                tuned.result}\SpecialCharTok{$}\NormalTok{RMSE)}
        \ControlFlowTok{break}
\NormalTok{      \}}

@@

\NormalTok{    param\_values.best\_result }\OtherTok{\textless{}{-}} 
      \FunctionTok{get\_best\_param.result}\NormalTok{(tuned.result}\SpecialCharTok{$}\NormalTok{parameter.value, }
\NormalTok{                          tuned.result}\SpecialCharTok{$}\NormalTok{RMSE)}
\NormalTok{  \}}
  \CommentTok{\# End repeat loop}
